Template Master's Thesis

### 2023-08-08 paper.tex
Update paper

\vfill
\begin{center}
%\includegraphics[width=4cm]{../figures/logo_cif.png}
\end{center}
\newpage
\singlespacing
%\vspace{-1.5cm}
\section*{Declaration of Independent Authorship}
I attest with my signature that I have written this work independently and without outside help. I also attest that the information concerning the sources used in this work is true and complete in every respect. All sources that have been quoted or paraphrased have been marked accordingly. 
Additionally, I affirm that any text passages written with the help of AI-supported technology are marked as such, including a reference to the AI-supported program used. This paper may be checked for plagiarism and use of AI-supported technology using the appropriate software. I understand that unethical conduct may lead to a grade of 1 or ``fail`` or expulsion from the study program.\\

@@

\begin{figure}[h!]
	\centering
	\hspace{-10cm}
	%\includegraphics[width=3cm]{../figures/signature.jpeg}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

@@

Public permissionless blockchains such as Bitcoin \citep{nakamotoBitcoin2008} and Ethereum \citep{buterin2014ethereum} allow users %individuals
 to participate with multiple pseudonymous addresses. The creation of these addresses is virtually cost free. Contrary to popular belief, these blockchains are entirely transparent. All transactions are stored as part of the blockchain's history and publicly observable.
This has opened up a nascent scientific field dealing with entity identification within blockchain networks. Researchers try to cluster addresses controlled by the same user by analyzing on-chain data and detecting usage patterns. The majority of these methods are heuristic which often makes them difficult to evaluate due to the absence of any ground truth. \newline
On one hand, identifying the addresses that belong to the same real-world entity is beneficial. According to \cite{FV:17} it allows better evaluation of network properties with respect to usage, distribution of wealth and detecting fraudulent activities. For instance, if a user distributes their voting power across various addresses, they might manipulate an on-chain voting process that seems fair at the outset. \newline
Conversely, the lack of privacy is also detrimental to most financial use cases (DeFi). As a response, multiple privacy-enhancing tools and protocols have been proposed to obfuscate transaction tracing. 
Nonetheless, these protocols are not yet widely adopted and careless user behaviour can undermine the privacy guarantees they offer. \newline
%As such/In conclusion
The topic of privacy, or lack thereof, will remain significant for future blockchain development and research.

@@

\textbf{Related work}\\
Previous work in this field can be broadly distinguished between heuristics developed for the unspent transaction output (UTXO) model (e.g, Bitcoin and ZCash) and the account-based model (e.g., Ethereum, Polygon). While both models share the concept of addresses, the notion of accounts is not present in UTXO-based blockchains. The way in which transactions are processed is fundamentally different, and therefore, clustering heuristics are not applicable to both paradigms.

A number of address clustering heuristics in the UTXO setup have been proposed for Bitcoin and derivatives \citep{Androulaki2013, Meiklejohn2013, Haslhofer2016, jourdan2018, kappos2022}. All these methods will not be the focus of this work and therefore not discussed. As suggested by Nakamoto in the Bitcoin whitepaper, most Bitcoin wallet implementations use a new key pair for each transaction to keep them from being linked to a common owner. \newline
In contrast to the UTXO-model, native transactions in account-based blockchains can only move funds between a single sender and receiver, and the ``change``remains in the sender account. Subsequent transactions necessarily use the same address again. The account-based model essentially relies on address reuse on the protocol level. Therefore, privacy guarantees should be lower and most users likely use only a small number of addresses. \citep{Beres2020} \newline
% Make an assumption that people only use a small set of addresses
Clustering heuristics for account-based blockchains were first introduced by \cite{FV:17}. He proposed and applied heuristics that exploit patterns related to deposit addresses, multiple participation in airdrops, and token authorization mechanisms. \newline
\cite{Beres2020} propose more universally applicable methods, as they argue that Victor’s heuristics, while powerful, assume participation in certain on-chain events. Their technique interprets transactions or token transfers as network graphs, with addresses as nodes and asset transfers as edges. They quantitavely compare graph-representation learning algorithms (a subset of machine learning) and propose further user profiling techniques based on time-of-day activity and transaction fees. Using ENS address pairs as ground truth, the authors rigorously test their methods and apply their findings to significantly reduce the privacy guarantees of Tornado Cash. \newline
\cite{wu2022tutela} extend on one of \cite{Beres2020} graph representation learning (node-embedding) algorithms and apply it at scale. Further, they also propose a set of new clustering heuristics targeting Tornado Cash, showing that careless user behavior can still reveal identity. They build an application built on those heuristics to measure the anonymity of an Ethereum address. All of the methods mentioned will be discussed in greater detail in Section 3. \newline
Broadening the scope of entity identification, \cite{victorlüders2019} study the largest Ethereum ERC-20 token networks from a graph perspective. Similarly, \cite{casalebrunet2021} apply a similar analysis to various non-fungible token (NFT) ecosystems. Both find that many of the networks follow either a star or a hub-and-spoke pattern. These patterns are similar to interaction graphs in social networks. Further, \cite{Payette2017} propose a segmentation of Ethereum address space into four distinct behaviour groups sharing similar attributes using k-means clustering. \newline
Rather than treating users as entities and analyzing on-chain data, \cite{yu2023} propose a novel approach for correlation analysis by exploiting network information. Although this approach has the potential to avoid the impact of privacy-enhancing technologies, it introduces a new set of limitations and problems. Thus, this approach may be of great interest for further research, particularly when privacy-enhancing techniques become more widely used.

\textbf{Our contribution}\\
In this work, we perform entity identification on an address set containing 473,927 addresses. To obtain these addresses, a separate project collected snapshots of avatar activity in Decentraland over an ten-month period from July 2022 to April 2023. Due to Decentraland's blockchain-based architecture, each avatar contains information about a user's Web3 address. 
We test the applicability of existing heuristics, and if feasible, their efficacy (in detecting entities using multiple addresses) is evaluated.
The main objective is to estimate the number of real-world entities that these addresses represent. The clustering of addresses is limited to this set only, without considering any (clusters for) addresses that were not registered/active in Decentraland during the given time frame.
Furthermore, we introduce our own heuristics for identifying address clusters and evaluate their suitability for implementation in other contexts.
%Anwenden/discuss and evaluate existing heuristics to my specific address set.
%Anpassen, erweitern

The remainder of this paper is structured as follows: In Section 2, we provide a brief overview of the basic concepts necessary to understand the setup. The include Ethereum accounts and the EVM, tokens, Decentraland, and privacy-enhancing protocols. Section 3 describes our methodology for data collection and preparation. The fourth section provides a detailed explanation of various clustering heuristics. Section 5 applies the clustering heuristics from 4. The fifth section applies the clustering heuristics described in section 4, after which we discuss the results of our analysis and summarize our findings in sections 5 and 6.

@@

\section{Preliminaries}

\textbf{Contract Accounts and EOAs}\\
Account-based blockchains usually distinguish between \textit{externally owned accounts} (EOAs) and \textit{contract accounts} (smart contracts). EOAs are created and controlled by private keys and can be used to hold the native protocol asset, send and receive transactions, and interact with contract accounts. Contract accounts are controlled by the contract's code, their state can be modified through transactions sent to the contract and they cannot initiate transactions. \citep{buterin2014ethereum} \newline Each account has a 20-byte address encoded in hexadecimal. For EOAs, this address is based on the last 20 bytes of the Keccak-256 hash of the ECDSA public key. For contract accounts, it is usually the last 20 bytes of the Keccak-256 hash of the RLP encoding of the sender address and account nonce. \citep{GW:14}

\textbf{Ethereum, Polygon and the EVM}\\
Ethereum and Polygon are both smart contract based account-model blockchains. They share the same execution logic - the Ethereum Virtual machine (EVM).
The EVM provides a standardized framework for contract execution, ensuring that smart contracts produce deterministic results across all nodes in the network \citep{GW:14}. The EVM is independent of the underlying blockchain protocol. This allows other blockchain implementations to adopt this standardized framework for contract execution.\newline
Polygon is a so-called sidechain and aims to address the scalability limitations of Ethereum. The adoption of the EVM means that both blockchains share the same user address schemes. It further enables developers to deploy and execute Ethereum-based smart contracts on the Polygon network with minimal changes. \citep{matic_whitepaper}
%A high-level programming language is typically used to write smart contracts, and a compiler to convert them into bytecode \citep{mastering_ethereum}. Solidity is currently the dominant smart contract language.
%Polygon PoS is a so-called sidechain, closely associated with Ethereum mainnet and used for scaling.
%Ethereum Mainnet is the chain on which much of the Decentraland infrastructure has been deployed.
%In addition, \cite{eip1014} introduces a new contract creation mechanism allowing for more predictable contract addresses.%EIP-1014
%This allows developers to calculate the address of a contract before it is deployed.

\textbf{Tokens and Token Standards}\\
According to \cite{roth2019tokenization}, \textit{tokens} as rivalrous, digital units of value that represent ownership of an asset or utility. Smart contracts are the primary method of creating tokens on EVM-based blockchains. In essence, a smart-contract-based token is a mapping of accounts with token balances and a set of functions defining how these balances can be changed. Any smart contract containing these elements can be interpreted as a token contract. \citep{roth2019tokenization} \newline
Token standards specify basic interfaces that allow for interoperability between contracts. These standards do not prescribe an implementation but set minimum requirements without restricting the design beyond that. 
The most common token standards are ERC-20\footnote{https://eips.ethereum.org/EIPS/eip-20} for fungible tokens (NFTs), ERC-721\footnote{https://eips.ethereum.org/EIPS/eip-721} for non-fungible tokens or ERC-1155\footnote{https://eips.ethereum.org/EIPS/eip-1155} for semi-fungible tokens.\newline
When a (blockchain) transaction completes, it produces a transaction receipt that contains log entries providing information about the actions that occurred during its execution \citep{mastering_ethereum}. The standarization of tokens allows us to listen for ``transfer events``, which are emitted whenever an asset/token changes ownership.
These transfer events include the \textit{sender} and \textit{recipient} address, along with a \textit{token ID} and/or \textit{amount}.

@@

\textbf{Decentraland} \\

%The term metaverse refers to Neal Stephenson’s 1992 novel Snow Crash and describes an immersive virtual world that is populated by humans-as-avatars.
Decentraland is the first large-scale blockchain-based virtual world. Users can own various digital assets, e.g., monetary units, land parcels, avatar apparel, or user names. The ownership of these assets is tracked and managed by smart contracts on Ethereum and Polygon. \citep{goldbergschaer}

 Why interesting? Business perspective

LAND / MANA

%If you choose to experience Decentraland Explorer without a wallet, the information will only be locally stored: you will be able to walk around, customize your Avatar and chat with others in-world, but you won’t have the chance to receive daily rewards, participate in events or log in with a different device using the same Guest ID and Avatar.

\textbf{Non-custodial crypto asset mixers and Tornado Cash} \\
%intuition of how it works,  distinction between custodial and non-custodial, Tornado cash = SC based
According to \cite{nadler2023tornado}, crypto-asset mixers are currently the most widely used approach to achieve privacy on public blockchains. At a high level, we can distinguish between custodial and non-custodial crypto-asset mixers. \newline
In the custodial setup, users send their assets to a centralized service provider's public deposit address. In return, the provider sends the assets back to a privately relayed recipient address. This approach is fully trust-based, as the service provider has complete control over the assets and access to the identifying data. \newline 
In contrast, non-custodial crypto asset mixers (e.g., Tornado Cash) replace the trusted mixing party with a publicly verifiable smart contract. They rely on cryptographic schemes that allow anyone to prove and independently verify the validity of a withdrawal without disclosing the link to a specific deposit. This has the advantage that users do not have to share the identifying information with anyone and there is no liquidity risk as the funds are locked. \citep{nadler2023tornado} \newline
In a nutshell, the mixer works in the following way: Various entities deposit the same amount of a specific crpyto asset to a mixer address acting as a pool. Anyone who has contributed to the pool may then generate a new address and withdraw their funds without revealing the link between the deposit and withdrawal addresses. This is achieved using zK-SNARKS. 
Each depositor inserts a hash value in a Merkle-tree. At the time of withdrawal, each legitimate withdrawer can prove unlinkably with a zero-knowledge proof that they know the pre-image of a previously inserted hash leaf in the Merkle-tree. %Subsequently, users can withdraw their asset from the mixer whenever they consider that the size of the anonymity set is satisfactory.
 For a more detailed description of how Tornado Cash works, see \cite{nadler2023tornado} or \cite{Beres2020}. For a mathematical discussion of ZKSNARK, please refer to Petkus (2019), Thaler (2022) or Berentsen et al. (2022). \newline
In the context of this work, it is sufficient to know that third parties can still observe the addresses that have deposited to and withdrawn from the pool. The number of addresses that have deposited in a particular pool is defined as the \textit{anonymity set}. Typically, a larger anonymity set will provide a higher level of privacy, as a larger number of deposit addresses can be associated with a given withdrawal address. \newline
In theory, if the anonymity set is large enough, third parties will not be able to link a specific depositor address to a specific withdrawal address. However, this is not always the case. For example, some users withdraw from the pool with the same address as they deposited. According to \cite{nadler2023tornado}, careless use of the protocol, testing transactions or external incentive mechanisms might be reasons for this behaviour. \newline
Furthermore, \cite{Beres2020} and \cite{wu2022tutela} have proposed approaches to de-anonymize Tornado cash transactions using various clustering heuristics. These methods are discussed in more detail in the next section.
% making use of address re-use, transaction activity, transaction cost choice, or more complex transaction graph and network analyses.

@@

\section{Data Collection and Preparation}

The original dataset comprised 473,927 distinct (Web3) addresses. As we only focus on clustering methods using on-chain transaction or token transfer data, we exclude addresses that have not been recorded on either Ethereum or Polygon. To accomplish this, we used data from Blockscan\footnote{\url{https://blockscan.com/}}. Overall, 59,651 addresses were recorded on Ethereum, 129,988 on Polygon, with 52,095 addresses appearing on both networks simultaneously. The set of remaining addresses consists of 137,544 addresses. Subsets are visualized in a Venn diagram in Figure 1. 
These addresses were used to collect transactions and token transfer events from Ethereum Mainnet and Polygon PoS. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider like Infura or Etherscan. The absence of account indexing in Ethereum or Polygon poses a challenge for retrieving all past transactions and token transfers of a specific address, as it requires scanning through all blocks and token transfer events emitted by designated token contracts. Fortunately, Etherscan offers/provides an API Endpoint Module for ``Accounts`` that facilitates the retrieval of transactions and token transfer events for a given address. By using our network-specific address subsets, we were able to significantly decrease the number of required API calls. 
We gathered all transaction and token transfer data of the addresses up until block 17,670,000 on Ethereum (July 11, 2023) and block 44,990,000 on Polygon (July 12, 2023). The output was saved in comma-separated value (CSV) files and imported into a MongoDB database\footnote{\url{https://www.mongodb.com/}}.

Statistics about data, how many transactions, how many transfers, size
Transfer events = 30,689,978
Transactions = 16,184,993, how many polygon/ethereum, table

\begin{table}[h]
\centering
\begin{tabular}{lcc}
\hline \hline
  & Transactions & Transfers \\
 \hline
 Polygon & Cell 2 & Cell 3 \\
 Ethereum & Cell 5 & Cell 6 \\
 \hline
 Total & 16,184,993 & 30,689,978 \\
 \hline \hline
\end{tabular}
\caption{Size of raw data}
\label{tab:data_size}
\end{table}

The transaction data fields contain the transaction hash, from and to address, the timestamp in seconds from the UNIX epoch, a nonce, value, input (often reffered to as calldata), the amount of gas used in the transaction, the specified gas price, the name of the function called (added by Etherscan) and the name of the blockchain.
Token transfer receipts include the transactions hash, timestamp, from and to address, nonce, gas price, the amount of gas used, chain information, the token contract address and name, tokenType, and token standard specific fields: value, tokenDecimal for ERC-20, tokenID for ERC-721, and value/tokenID for ERC-1155.

@@

@@

\section{Clustering Heuristics}
% Goal: explain each heuristic, categorize, and discuss applicability to our address set

\textbf{Self-authorization} \newline
The \texttt{approve} function required in all three token standards allows another address to spend tokens on behalf of the actual owners. This functionality is mainly used in connection with smart contracts, but can also be applied for regular EOAs. Users could approve another address they own, a reason for this might be risk distribution over several addresses with partial accessibility. \citep{FV:17}

\textbf{Deposit address reuse} \newline
This heuristic exploits the common practice of crypto exchanges creating so-called deposit addresses for each user, which forward funds to a main address. Multiple addresses that send funds to the same deposit address are highly likely controlled by the same entity. \citep{FV:17}

\textbf{Airdrop multi participation} \newline
Airdrops are popular token distribution mechanisms, and recipients are mostly chosen based on past protocol activity. A famous example is Uniswap's UNI airdrop, where each address that interacted with the protocol received a fixed amount of UNI tokens\footnote{\url{https://blog.uniswap.org/uni}}. These distribution mechanisms are often not Sybil-resistant, leading to people creating multiple addresses in anticipation of an airdrop. However, managing the tokens on all of these addresses is impractical, which is why they are often aggregated into one address. \citep{FV:17}

>>>>

\vfill
\begin{center}
\includegraphics[width=4cm]{../figures/logo_cif.png}
\end{center}
\newpage
\singlespacing
%\vspace{-1.5cm}
\section*{Declaration of Independent Authorship}
I attest with my signature that I have written this work independently and without outside help. I also attest that the information concerning the sources used in this work is true and complete in every respect. All sources that have been quoted or paraphrased have been marked accordingly. 
Additionally, I affirm that any text passages written with the help of AI-supported technology are marked as such, including a reference to the AI-supported program used. This paper may be checked for plagiarism and use of AI-supported technology using the appropriate software. I understand that unethical conduct may lead to a grade of 1 or ``fail`` or expulsion from the study program.\\

@@

\begin{figure}[h!]
	\centering
	\hspace{-10cm}
	\includegraphics[width=3cm]{../figures/signature.jpeg}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

@@

Public permissionless blockchains such as Bitcoin \citep{nakamotoBitcoin2008} and Ethereum \citep{buterin2014ethereum} allow users %individuals
 to participate with multiple pseudonymous addresses. The creation of these addresses is virtually cost free. Contrary to popular belief, these blockchains are entirely transparent. All transactions are stored as part of the blockchain's history and publicly observable.
This has opened up a nascent scientific field dealing with entity identification within blockchain networks. Researchers try to cluster addresses controlled by the same user by analyzing on-chain data and detecting usage patterns. The frequent lack of ground truth makes it difficult to evaluate different clustering methods. As a result, most methods are heuristic. %The majority of these methods are heuristic which often makes them difficult to evaluate due to the absence of any ground truth.
\newline On one hand, identifying the addresses that belong to the same real-world entity is beneficial. According to \cite{FV:17} it allows better evaluation of network properties with respect to usage, distribution of wealth and detecting fraudulent activities. For instance, if a user distributes their voting power across various addresses, they might manipulate an on-chain voting process that seems fair at the outset. \newline
Conversely, the lack of privacy is also detrimental to most financial use cases (DeFi). As a response, multiple privacy-enhancing tools and protocols have been proposed to obfuscate transaction tracing. 
Nonetheless, these protocols are not yet widely adopted and careless user behaviour can undermine the privacy guarantees they offer. \newline
%As such/In conclusion
The topic of privacy, or lack thereof, will remain significant for future blockchain development and research.

@@

\textbf{Related work}\\
Previous work in this field can be broadly distinguished between methods%heuristics
 developed for the unspent transaction output (UTXO) model (e.g, Bitcoin and ZCash) and the account-based model (e.g., Ethereum, Polygon). While both models share the concept of addresses, the notion of accounts is not present in UTXO-based blockchains. The way in which transactions are processed is fundamentally different, and therefore, clustering heuristics are not applicable to both paradigms.

A number of address clustering heuristics in the UTXO setup have been proposed for Bitcoin and derivatives \citep{Androulaki2013, Meiklejohn2013, Haslhofer2016, jourdan2018, kappos2022}. All these methods will not be the focus of this work and therefore not discussed. As suggested by \cite{nakamotoBitcoin2008} in the Bitcoin whitepaper, most Bitcoin wallet implementations use a new key pair for each transaction to keep them from being linked to a common owner. \newline
In contrast to the UTXO-model, native transactions in account-based blockchains can only move funds between a single sender and receiver, and the ``change``remains in the sender account. Subsequent transactions necessarily use the same address again. The account-based model essentially relies on address reuse on the protocol level \citep{Beres2020}. Therefore, privacy guarantees should be lower and most users likely use only a small number of addresses. \newline
% Make an assumption that people only use a small set of addresses
Clustering heuristics for account-based blockchains were first introduced by \cite{FV:17}. He proposed and applied heuristics that exploit patterns related to deposit addresses, multiple participation in airdrops, and token authorization mechanisms. \newline
\cite{Beres2020} propose more universally applicable methods, as they argue that Victor’s heuristics, while powerful, assume participation in certain on-chain events. Their technique interprets transactions or token transfers as network graphs, with addresses as nodes and asset transfers as edges. They quantitavely compare graph-representation learning algorithms (a subset of machine learning) and propose further user profiling techniques based on time-of-day activity and transaction fees. Using ENS address pairs as ground truth, the authors rigorously test their methods and apply their findings to significantly reduce the privacy guarantees of Tornado Cash, a non-custodial privacy-enhancing protocol on Ethereum. \newline
\cite{wu2022tutela} extend on one of \cite{Beres2020} graph representation learning %node-embedding
 algorithms and apply it at scale. Further, they also propose a set of new clustering heuristics targeting Tornado Cash,  showing that careless user behavior can still reveal identity. They build an application built on those heuristics to measure the anonymity of an Ethereum address. All of the methods mentioned will be discussed in greater detail in Section 3. \newline
Broadening the scope of entity identification, \cite{victorlüders2019} study the largest Ethereum ERC-20 token networks from a graph perspective. Similarly, \cite{casalebrunet2021} apply network graph analysis to various non-fungible token (NFT) ecosystems. Both find that many of the networks follow either a star or a hub-and-spoke pattern. These patterns are common to interaction graphs in social networks. Further, \cite{Payette2017} propose a segmentation of the Ethereum address space into four distinct behaviour groups sharing similar attributes using k-means clustering. \newline
Rather than treating users as entities and analyzing on-chain data, \cite{yu2023} propose a novel approach for correlation analysis by exploiting network information. Although this approach has the potential to avoid the impact of privacy-enhancing technologies, it introduces a new set of limitations and problems. Thus, this approach may be of great interest for further research, particularly when privacy-enhancing techniques become more widely used.

\textbf{Our contribution}\\
In this work, we perform entity identification on an address set containing 473,927 addresses. To obtain these addresses, a separate project collected snapshots of avatar activity in Decentraland over an ten-month period from July 2022 to April 2023. Due to Decentraland's blockchain-based architecture, each avatar contains information about a user's Web3 address. 
We test the applicability of existing heuristics, and if feasible, their efficacy in detecting entities using multiple addresses is evaluated.
The main objective is to estimate the number of real-world entities that these addresses represent. The clustering of addresses is limited to this set only, without considering any %(clusters for)
 addresses outside of it, e.g., that were not registered in Decentraland during the given time frame.
Furthermore, we introduce our own heuristics for identifying address clusters and evaluate their suitability for implementation in other contexts.
%Anwenden/discuss and evaluate existing heuristics to my specific address set.
%Anpassen, erweitern

The remainder of this paper is structured as follows: In Section 2, we provide a brief overview of the basic concepts necessary to understand the setup. The include Ethereum accounts and the EVM, tokens, Decentraland, and privacy-enhancing protocols. Section 3 describes our methodology for data collection and preparation. The fourth section provides a detailed explanation of various clustering heuristics.The fifth section applies the clustering heuristics described in section 4, after which we discuss the results of our analysis and summarize our findings in sections 6 and 7.

@@

\section{Preliminaries}
% Introduce section
\subsection{Contract Accounts and EOAs}
Account-based blockchains usually distinguish between \textit{externally owned accounts} (EOAs) and \textit{contract accounts} (smart contracts). EOAs are created and controlled by private keys and can be used to hold the native protocol asset, send and receive transactions, and interact with contract accounts. Contract accounts are controlled by the contract's code, their state can be modified through transactions sent to the contract and they cannot initiate transactions. \citep{buterin2014ethereum} \newline Each account has a 20-byte address encoded in hexadecimal. For EOAs, this address is based on the last 20 bytes of the Keccak-256 hash of the ECDSA public key. For contract accounts, it is usually the last 20 bytes of the Keccak-256 hash of the RLP encoding of the sender address and account nonce. \citep{GW:14}

\subsection{Ethereum, Polygon and the EVM}
Ethereum and Polygon are both smart contract based account-model blockchains. They share the same execution logic - the Ethereum Virtual machine (EVM).
The EVM provides a standardized framework for contract execution, ensuring that smart contracts produce deterministic results across all nodes in the network \citep{GW:14}. The EVM is independent of the underlying blockchain protocol. This allows other blockchain implementations to adopt this standardized framework for contract execution.\newline
Polygon is a so-called sidechain and aims to address the scalability limitations of Ethereum. The adoption of the EVM means that both blockchains share the same user address schemes. It further enables developers to deploy and execute Ethereum-based smart contracts on the Polygon network with minimal changes. \citep{matic_whitepaper}
%A high-level programming language is typically used to write smart contracts, and a compiler to convert them into bytecode \citep{mastering_ethereum}. Solidity is currently the dominant smart contract language.
%Polygon PoS is a so-called sidechain, closely associated with Ethereum mainnet and used for scaling.
%Ethereum Mainnet is the chain on which much of the Decentraland infrastructure has been deployed.
%In addition, \cite{eip1014} introduces a new contract creation mechanism allowing for more predictable contract addresses.%EIP-1014
%This allows developers to calculate the address of a contract before it is deployed.

\subsection{Tokens and Token Standards}
\cite{roth2019tokenization} define \textit{tokens} as rivalrous, digital units of value that represent ownership of an asset or utility. Smart contracts are the primary method of creating tokens on EVM-based blockchains. In essence, a smart-contract-based token is a mapping of accounts with token balances and a set of functions defining how these balances can be changed. Any smart contract containing these elements can be interpreted as a token contract. \citep{roth2019tokenization} \newline
Token standards specify basic interfaces that allow for interoperability between contracts. These standards do not prescribe an implementation but set minimum requirements without restricting the design beyond that \citep{mastering_ethereum}. 
The most common token standards are ERC-20\footnote{https://eips.ethereum.org/EIPS/eip-20} for fungible tokens, ERC-721\footnote{https://eips.ethereum.org/EIPS/eip-721} for non-fungible tokens (NFTs) and ERC-1155\footnote{https://eips.ethereum.org/EIPS/eip-1155} for semi-fungible tokens.\newline
When a (blockchain) transaction completes, it produces a transaction receipt that contains log entries providing information about the actions that occurred during its execution \citep{mastering_ethereum}. The standarization of tokens allows us to listen for ``transfer events``, which are emitted whenever an token changes ownership.
These transfer events include the \textit{sender} and \textit{recipient} address, along with a \textit{token ID} and/or \textit{amount}.

@@

\subsection{Decentraland}
\textit{Decentraland} is the first large-scale blockchain-based virtual world. It has attracted the attention of well-known companies like Tommy Hilfiger, Samsung, PepsiCo, Diesel, Adidas, and Netflix, who are actively engaged in it  \citep{metaverse-retailing2023}.\newline
Decentraland is an ideal platform for empirical research for two main reasons.
\textit{First}, its open architecture allows compiling snapshots of user activity, including the users' locations  within the metaverse. This was done by \cite{metaverse-retailing2023}, who captured snapshots of users/avatar locations over a period of 10 months. The addresses used in this paper are provided form this work.
\textit{Second}, users who connect their Web3 account can own various digital assets, such as monetary units, land parcels, and avatar collectibles (e.g. wearables, emotes, and names). Smart contracts on Ethereum and Polygon track and manage the ownership of these assets \cite{goldbergschaer2023}. As a result, this allows us to analyze the entire transaction history and derive information about the avatars’/users' economic activity.  \newline 
The Decentraland ecosystem has two main tokens. \textit{LAND}, an NFT compliant with ERC-721, manages the ownership of land parcels in Decentraland. \textit{MANA}, a fungible token compliant with ERC-20, is the in-world currency used to purchase digital goods and services. The majority of purchases and trades are settled in MANA. Furthermore, Decentraland %the foundation/DAO/developers
operates/provides/hosts a virtual/smart-contract-based marketplace where its users may buy or exchange LAND or other in-game collectibles.

%Owners of multiple adjacent land parcels can combine them into an ESTATE, which acts as a wrapper token for the land parcels.
%The term metaverse refers to Neal Stephenson’s 1992 novel Snow Crash and describes an immersive virtual world that is populated by humans-as-avatars.
%Goods: in-world assets, such as avatar wearables, merchandise, or tickets, as well as tokenized vouchers that can be redeemed for real-world goods or services.

\subsection{Non-custodial crypto asset mixers}
%intuition of how it works,  distinction between custodial and non-custodial, Tornado cash = SC based
According to \cite{nadler2023tornado}, crypto-asset mixers are currently the most widely used approach to achieve privacy on public blockchains. At a high level, we can distinguish between custodial and non-custodial crypto-asset mixers. \newline
In the custodial setup, users send their assets to a centralized service provider's public deposit address. In return, the provider sends the assets back to a privately relayed recipient address. This approach is fully trust-based, as the service provider has complete control over the assets and access to the identifying data. \newline 
In contrast, non-custodial crypto asset mixers (e.g., Tornado Cash) replace the trusted mixing party with a publicly verifiable smart contract. They rely on cryptographic schemes that allow anyone to prove and independently verify the validity of a withdrawal without disclosing the link to a specific deposit. This has the advantage that users do not have to share the identifying information with anyone and there is no liquidity risk as the funds are locked. \citep{nadler2023tornado} \newline
In a nutshell, the mixer works in the following way: Various entities deposit the same amount of a specific crpyto asset to a mixer address acting as a pool. Anyone who has contributed to the pool may then generate a new address and withdraw their funds without revealing the link between the deposit and withdrawal addresses. This is achieved using zK-SNARKS. 
Each depositor inserts a hash value in a Merkle-tree. At the time of withdrawal, each legitimate withdrawer can prove unlinkably with a zero-knowledge proof that they know the pre-image of a previously inserted hash leaf in the Merkle-tree. %Subsequently, users can withdraw their asset from the mixer whenever they consider that the size of the anonymity set is satisfactory.
 For a more detailed description of how Tornado Cash works, see \cite{nadler2023tornado} or \cite{Beres2020}. For a mathematical discussion of ZKSNARKs, please refer to Petkus (2019), Thaler (2022) or Berentsen et al. (2022). \newline
In the context of this work, it is sufficient to know that third parties can still observe the addresses that have deposited to and withdrawn from the pool. The number of addresses that have deposited in a particular pool is defined as the \textit{anonymity set}. Typically, a larger anonymity set will provide a higher level of privacy, as a more deposit addresses can be associated with any given withdrawal address. \newline
In theory, if the anonymity set is large enough, third parties will not be able to link a specific depositor address to a specific withdrawal address. However, this is not always the case. For example, some users withdraw from the pool with the same address as they deposited. According to \cite{nadler2023tornado}, careless use of the protocol, testing transactions or external incentive mechanisms might be reasons for this behaviour. \newline
Furthermore, \cite{Beres2020} and \cite{wu2022tutela} have proposed approaches to de-anonymize Tornado cash transactions using various clustering heuristics. These methods are discussed in more detail in the next section.
% making use of address re-use, transaction activity, transaction cost choice, or more complex transaction graph and network analyses.

@@

\section{Data Collection and Preparation}
% Introduction Satz - Ausgangslage 
The starting point for our work was a dataset comprising 473,927 distinct (Web3) addresses. %The original dataset comprised 473,927 distinct (Web3) addresses.
 As we focus on clustering methods using on-chain data, we exclude addresses that have not been recorded on either Ethereum or Polygon. To accomplish this, we used data from Blockscan\footnote{\url{https://blockscan.com/}}. Overall, 59,651 addresses were recorded on Ethereum, 129,988 on Polygon, with 52,095 addresses appearing on both networks simultaneously. The set of remaining addresses consists of 137,544 addresses. Subsets are visualized as a Venn diagram in \ref{fig:Venn}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.6\textwidth]{../figures/venn_diagram.png}
	\caption{Venn Diagram of address subsets}
	\label{fig:Venn}
\end{figure}

\subsection{Transactions and Token Transfer Events}
The address dataset was used to collect transaction and token transfer event data from Ethereum and Polygon. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider like Infura or Etherscan. The absence of account indexing in Ethereum or Polygon poses a challenge for retrieving all past transactions and token transfers of a specific address, as it requires scanning through all blocks and token transfer events emitted by designated token contracts. Fortunately, Etherscan offers an API Endpoint Module for ``Accounts`` that facilitates the retrieval of transactions and token transfer events for a given address. By using our network-specific address subsets, we were able to significantly decrease the number of required API calls. 
We gathered all transactions and token transfers up until block height \texttt{17,670,000} on Ethereum (July 11, 2023) and block \texttt{44,990,000} on Polygon (July 12, 2023). The output was saved in separate CSV files. We then imported this data into a MongoDB database\footnote{\url{https://www.mongodb.com/}}.
In total, we collected 30,689,978 token transfers and 16,092,531 normal transactions. Figure X visualizes the number of daily transactions and token transfers for each chain. Can see activity.

Transfer Events: 30,689,978
Transfer Events Ethereum: 7,832,778
Transfer Events Polygon: 22,857,200

Statistics about data, how many transactions on Polygon, how many on Ethereum, how many transfers on Polygon, how many on Ethereum -> Show in Figure

Transactions = 16,092,531
Transactions Ethereum = 8,448,584
Transactions Polygon = 7,643,947

%\begin{table}[h]
%\centering
%\begin{tabular}{lcc}
%\hline \hline
%  & Transactions & Transfers \\
% \hline
% Polygon & Cell 2 & Cell 3 \\
% Ethereum & Cell 5 & Cell 6 \\
% \hline
% Total & 16,184,993 & 30,689,978 \\
% \hline \hline
%\end{tabular}
%\caption{Size of raw data}
%\label{tab:data_size}
%\end{table}

Transaction hash, timestamp in seconds from the UNIX epoch, gas used in the transaction, specified gas price, nonce

The transaction data fields contain the transaction hash, from and to address, the timestamp in seconds from the UNIX epoch, a nonce, value, input (often referred to as calldata), the amount of gas used in the transaction, the specified gas price, the name of the function called (added by Etherscan) and the name of the blockchain.

Token transfer receipts include the transaction hash, UNIX timestamp, sender and recipient address, nonce, gas price, the amount of gas, chain information, the token contract address and name, tokenType, and token standard specific fields: value, tokenDecimal for ERC-20, tokenID for ERC-721, and value/tokenID for ERC-1155.

@@

\subsection{ENS Names}

@@

\section{Clustering Heuristics}
In this section, we provide a overview of clustering heuristics in the context of account-based blockchains and on-chain data. We will give a detailed explanation of the intuition/rationale behind the heuristic, the workings and discuss the applicability for our case. In Section 5 we will then apply and, if possible, evaluate the heuristics deemed suited for us.

%detailed explanation
% Goal: explain each heuristic, categorize, and discuss applicability to our address set

\subsection{Self-authorization}
The \texttt{approve} function required in all three token standards allows another address to spend tokens on behalf of the actual owners. This functionality is mainly used in connection with smart contracts, but can also be applied for regular EOAs. Users could approve another address they own, a reason for this might be risk distribution over several addresses with partial accessibility. \citep{FV:17}

\subsection{Deposit address reuse}
This heuristic exploits the common practice of crypto exchanges creating so-called deposit addresses for each user, which forward funds to a main address. Multiple addresses that send funds to the same deposit address are highly likely controlled by the same entity. \citep{FV:17}
